Bayes Optimal Informer Sets for Drug Discovery\thanks{abc}

### 2020-04-17 BOISE/Paper_draft/draft.tex
Add Proposition 1

Theorem 1 claims that for any given informer $A$ and intermediate data $x_A$, there exists a top-set rule 
\begin{eqnarray*}
T^*(A,x_A,x_0) = \left\{ j \in J: \; \sum_{k\in J} \mathbbm{1}(\hat \theta_{i^*k} > \hat \theta_{i^*j}) \leq n_T \right\}
\end{eqnarray*}
such that  $r(A,T) \geq r(A, T^*)$. In other words, given $A$ and $x_A$, the best top-set rule is to select top $n_T$ compounds with largest posterior mean. Recall the expression of Bayes risk: 
\begin{align*}
r(A, T) 
&= \int_{x_0} \int_{x_A} \int_\Theta  L\left( A, T \right) \, 
p(\theta, x_A,x_0)\, d\theta \, d x_A \, d x_0\\
&= \int_{x_0} \int_{x_A} \int_\Theta  L\left( A, T \right) \, 
p(\theta |x_A,x_0)p(x_A|x_0)p(x_0)\, d\theta \, d x_A \, d x_0
\end{align*}
To prove the Theorem, it is sufficient to show that for any possible fixed $x_A$, $x_0$, inequality 
\begin{eqnarray}
\label{eq:inequal1}
\int_\Theta  L\left( A, T \right) \, 
p(\theta |x_A,x_0) \, d\theta \geq \int_\Theta  L\left( A, T^* \right) \, 
p(\theta |x_A,x_0) \, d\theta
\end{eqnarray}holds, and this is obvious if we write loss function $L(A,T)$ explicitly:
\begin{align*}
\int_\Theta  L\left( A, T \right) \, p(\theta |x_A,x_0) \, d\theta &= \int_\Theta  ( n_T - \sum_{j\in T} \theta_{i^*j}) 
p(\theta |x_A,x_0) \, d\theta\\
&= n_T - \sum_{j\in T} \hat \theta_{i^*j}\\
&\geq n_T - \sum_{j\in T^*}  \hat \theta_{i^*j}\\
&= \int_\Theta  L\left( A, T^* \right) \, 
p(\theta |x_A,x_0) \, d\theta
\end{align*}
The equality comes from previous definition that $\hat \theta_{i^*j} = \mathbb E(\theta_{i^*j}|x_A,x_0)$.
\end{proof}

\subsection{Proof of Theorem 2}
\begin{proof}

@@

For simplicity of notation, we denote top set rule as 
$T_1^*(x_A,x_k) = T^*(A\cup\{k\}, x_{A\cup\{k\}}, x_0)$ and $T_2^*(x_A) = T^*(A, x_{A}, x_0)$. Recall that:
\begin{eqnarray*}
    {\mbox {\rm PEL}}_1(A\cup\{k\})= \int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right) 
\left[ \sum_{j \in  T_1^*(x_A,x_k)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]dx_A dx_k
\end{eqnarray*}

@@

\begin{align*}
&\int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right)
\left[ \sum_{j \in  T_1^*(x_A,x_k)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]\\ 
&\leq 
\int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right)
\left[ \sum_{j \in  T_2^*(x_A)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]
\end{align*}

Notice that in our model, $x_A \indep x_k \, | \, \theta_{i^*}$ and $\theta_{i^*A}\indep \theta_{i^*k}\, | \, x_0$. So we have $x_A\indep x_k\,|\, x_0$, and hence:

\begin{align*}
&\int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right)
\left[ \sum_{j \in  T_2^*(x_A)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]\\
    &= \int_{x_{A}}p\left(x_A|x_0\right)\sum_{j \in  T_2^*(x_A)}\left[\int_{x_k}p(x_k|x_0)\left(1-\mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k)\right)\right]\\
&= \int_{x_{A}}p\left(x_A|x_0\right) \sum_{j \in  T_2^*(x_A)}\left[  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A) \right].
\end{align*}

>>>>

Theorem 1 claims that for any given informer $A$ and intermediate data $x_A$, there exists a top-set rule 
\begin{eqnarray*}
T^*(A,x_A,x_0) = \left\{ j \in J: \; \sum_{k\in J} \mathbbm{1}(\hat \theta_{i^*k} > \hat \theta_{i^*j}) \leq n_T \right\}
\end{eqnarray*}
such that  $r(A,T) \geq r(A, T^*)$. In other words, given $A$ and $x_A$, the best top-set rule is to select top $n_T$ compounds with largest posterior mean. Recall the expression of Bayes risk: 
\begin{align*}
r(A, T) 
&= \int_{x_0} \int_{x_A} \int_\Theta  L\left( A, T \right) \, 
p(\theta, x_A,x_0)\, {\rm d}\theta \, {\rm d} x_A \, {\rm d} x_0\\
&= \int_{x_0} \int_{x_A} \int_\Theta  L\left( A, T \right) \, 
p(\theta |x_A,x_0)p(x_A|x_0)p(x_0)\, {\rm d}\theta \, {\rm d} x_A \, {\rm d} x_0
\end{align*}
To prove the Theorem, it is sufficient to show that for any possible fixed $x_A$, $x_0$, inequality 
\begin{eqnarray}
\label{eq:inequal1}
\int_\Theta  L\left( A, T \right) \, 
p(\theta |x_A,x_0) \, {\rm d}\theta \geq \int_\Theta  L\left( A, T^* \right) \, 
p(\theta |x_A,x_0) \, {\rm d}\theta
\end{eqnarray}holds, and this is obvious if we write loss function $L(A,T)$ explicitly:
\begin{align*}
\int_\Theta  L\left( A, T \right) \, p(\theta |x_A,x_0) \, {\rm d}\theta &= \int_\Theta  ( n_T - \sum_{j\in T} \theta_{i^*j}) 
p(\theta |x_A,x_0) \, {\rm d}\theta\\
&= n_T - \sum_{j\in T} \hat \theta_{i^*j}\\
&\geq n_T - \sum_{j\in T^*}  \hat \theta_{i^*j}\\
&= \int_\Theta  L\left( A, T^* \right) \, 
p(\theta |x_A,x_0) \, {\rm d}\theta
\end{align*}
The equality comes from previous definition that $\hat \theta_{i^*j} = \mathbb E(\theta_{i^*j}|x_A,x_0)$.
\end{proof}

\subsection{Proof of Theorem 2}
We first prove a lemma that will be used in proof of Theorem 2:
\begin{proposition}
In BOISE model (\ref{eq:model}), if we use $x_A$ and $x_k$ to denote subsets of outcomes on the new target: $x_{i^*A}$ and $x_{i^*k}$, and $x_{0A}$ and $x_{0k}$ to denote subsets of outcomes on the original data set $x_0$, then we have:
\begin{eqnarray}
\label{condition_ind}
x_A \indep x_k \, | \, x_0.
\end{eqnarray}
\end{proposition}

\begin{proof}[Proof of Proposition 1]
Notice that in model (\ref{eq:model}) we have:
\begin{align}
\label{condition_ind2}
\begin{split}
    (x_A,x_k) &\indep x_0 \, | \, (\theta_{i^*A},\theta_{i^*k});\\
x_A &\indep x_k \,|\, (\theta_{i^*A},\theta_{i^*k});\\
x_{0A} &\indep x_{0k} \,|\, (\theta_{i^*A},\theta_{i^*k});\\
\theta_{i^*A} &\indep \theta_{i^*k};\\
x_{0A} &\indep x_{0k}.
\end{split}
\end{align}
From independence and conditional independence in (\ref{condition_ind2}), we can show that $\theta_{i^*A}\indep \theta_{i^*k}\, | \, x_0$ as following:

\begin{align*}
    p(\theta_{i^*A},\theta_{i^*k}\,|\,x_0)&=
    \frac{p(x_0\,|\,\theta_{i^*A},\theta_{i^*k})p(\theta_{i^*A},\theta_{i^*k})}{p(x_0)}\\
    &= \frac{p(x_{0A},x_{0k}\,|\,\theta_{i^*A},\theta_{i^*k})p(\theta_{i^*A},\theta_{i^*k})}{p(x_{0A},x_{0k})}\\
    &= p(\theta_{i^*A}\,|\,x_{0A})p(\theta_{i^*k}\,|\,x_{0k})\\
    &= p(\theta_{i^*A}\,|\,x_0)p(\theta_{i^*k}\,|\,x_0).
\end{align*}

Therefore, for desired $p(x_A,x_k\,|\,x_0)$ we have:
\begin{align}
\begin{split}
    p(x_A,x_k\,|\,x_0) &= \int_\Theta p(x_A,x_k\,|\,\theta_{i^*A},\theta_{i^*k},x_0)p(\theta_{i^*A},\theta_{i^*k}\,|\,x_0) {\rm d}\theta_{i^*A}{\rm d}\theta_{i^*k}\\
    &= \int_\Theta p(x_A\,|\,\theta_{i^*A})p(x_k\,|\,\theta_{i^*k})p(\theta_{i^*A}\,|\,x_0)p(\theta_{i^*k}\,|\,x_0){\rm d}\theta_{i^*A}{\rm d}\theta_{i^*k}\\
    &= p(x_A\,|\,x_0)p(x_k\,|\,x_0).
\end{split}
\end{align}
and this is equivalent to $x_A \indep x_k \, | \, x_0$.
\end{proof}

With the help of Proposition 1 we can now prove Theorem 2:

\begin{proof}[Proof of Theorem 2]

@@

For simplicity of notation, we denote top set rule as 
$T_1^*(x_A,x_k) = T^*(A\cup\{k\}, x_{A\cup\{k\}}, x_0)$ and $T_2^*(x_A) = T^*(A, x_{A}, x_0)$. Recall that:
\begin{eqnarray*}
    {\mbox {\rm PEL}}_1(A\cup\{k\})= \int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right) 
\left[ \sum_{j \in  T_1^*(x_A,x_k)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]{\rm d}x_A {\rm d}x_k
\end{eqnarray*}
Here we still use $x_A$ and $x_k$ to denote $x_{i^*A}$ and $x_{i^*k}$, a subset of outcomes on the new target $i^*$.

@@

\begin{align*}
&\int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right)
\left[ \sum_{j \in  T_1^*(x_A,x_k)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right] {\rm d}x_A {\rm d}x_k\\
&\leq 
\int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right)
\left[ \sum_{j \in  T_2^*(x_A)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]{\rm d}x_A {\rm d}x_k
\end{align*}

From Proposition 1, $x_A \indep x_k \, | \, x_0$, and hence we have:

\begin{align*}
&\int_{x_{A}}\int_{x_k} p\left( x_{A},x_k | x_0 \right)
\left[ \sum_{j \in  T_2^*(x_A)}  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k) \right]{\rm d}x_A {\rm d}x_k\\
    &= \int_{x_{A}}p\left(x_A|x_0\right)\sum_{j \in  T_2^*(x_A)}\left[\int_{x_k}p(x_k|x_0)\left(1-\mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A,x_k)\right){\rm d}x_k\right] {\rm d}x_A\\
&= \int_{x_{A}}p\left(x_A|x_0\right) \sum_{j \in  T_2^*(x_A)}\left[  1 - \mathbb{E} (\theta_{i^*,j} \, |\, x_0,x_A) \right]{\rm d}x_A.
\end{align*}
